---
title: "Practical Machine Learning Course Project"
author: "Avinash Singh Pundhir"
date: "December 21, 2015"
output: html_document
---

Human Activity Recognition (HAR) is a new key field of study that is emerging as a key research area to develop context aware systems. There are many potential applications of HAR. Some of those are following:

* Elderly monitoring 
* Life log systems for monitoring energy expenditure and for supporting weight-loss programs
* Digital assistants for weight lifting exercises

Weight Lifting Exercises Dataset:

Weight Lifting Exercises dataset is to investigate "how (well)" an activity was performed by the wearer. Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3v3mADg7n

```{r, echo=TRUE, message=FALSE}
library(caret)
library(dplyr)
```

### Getting and Cleaning Data

```{r}
setwd("C:\\Users\\Avinash\\Desktop\\R Files\\Coursera\\Machine Learning")

hardata <- read.csv("pml-training.csv")

#Cleaning Data 
hardata <- select(hardata, 
                  -X, 
                  -user_name, 
                  -raw_timestamp_part_1, 
                  -raw_timestamp_part_2, 
                  -cvtd_timestamp, 
                  -new_window, 
                  -num_window)
#Extracting belt, arm, dumbell, forearm related variables only
hardata <- hardata[, grepl("gyros|accel|magnet|classe", names(hardata))]
hardata <- select(hardata, -starts_with("var"))

dim(hardata)
```

### Creating Training and Test Sets

```{r}
set.seed(1456)
intrain <- createDataPartition(y = hardata$classe, 
                               p = 0.6, 
                               list = FALSE)
traincomplete <- hardata[intrain,]
testing <- hardata[-intrain,]
dim(traincomplete)
dim(testing)
```

### Feature Plots for Exploratory Analysis

```{r}
featurePlot(x = traincomplete[, c("total_accel_belt", 
                                  "total_accel_arm", 
                                  "total_accel_dumbbell",
                                  "total_accel_forearm")],
            y = traincomplete$classe,
            plot = "ellipse",
            auto.key = list(columns = 4))
```

### Fitting Random Forest Classification Model

```{r, echo=TRUE, message=FALSE, cache=TRUE}

#Adding Repeated K Fold Cross Validation

fitCOntrol <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 3)

rffit <- train(classe ~ ., 
               data = traincomplete, 
               method = "rf",
               trControl = fitCOntrol)

print(rffit, digits = 3)
print(rffit$finalModel)

pred1 <- predict(rffit, testing)

confusionMatrix(testing$classe, pred1)

plot(rffit, 
     log = "y", 
     lwd = 2, 
     main = "Random forest accuracy", 
     xlab = "Predictors", 
     ylab = "Accuracy")
```

* Here we have used repeated K fold cross validation approach for cross validation. Where we have taken K as 10. As larger K would give us less bias and more variance. While using smaller K i.e. in our case we will get more bias and less variance.

### Out of Sample Error

```{r}
oosaccuracy <- sum(pred1 == testing$classe)/length(pred1)
oose <- round((1 - oosaccuracy)*100,3)
```

* Out of Sample error estimation is `r oose` 

### Prediction for 20 Test Cases

```{r, echo=TRUE}
test <- read.csv("pml-testing.csv")
test <- rename(test, classe = problem_id)
test <- select(test, 
                  -X, 
                  -user_name, 
                  -raw_timestamp_part_1, 
                  -raw_timestamp_part_2, 
                  -cvtd_timestamp, 
                  -new_window, 
                  -num_window)
#Extracting belt, arm, dumbell, forearm related variables only
test <- test[, grepl("gyros|accel|magnet|classe", names(test))]
test <- select(test, -starts_with("var"))
pred2 <- predict(rffit, test)
pred2
```
